---
title: "Data607_Assignment03"
author: "Ryan Weber"
date: "February 18, 2018"
output: 
  html_document:
    theme: lumen
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
library(stringr)
library(dplyr)
library(tidyr)
```

Github for homework: https://github.com/rweberc/Data607_Project1

  
### Read in data  
```{r}
con <- file("https://raw.githubusercontent.com/rweberc/Data607_Project1/master/rawdata.txt", open="r")
line <- readLines(con)
close(con)
```  
  
  
### Remove dashed lines 
```{r}
line <- line[!str_detect(line, "^-")]
line[1:5]
```

  
  
### Create data.frame with a single column out of the indidivual 'lines'
```{r}
rawData <- data.frame(col1 = line[3:length(line)], stringsAsFactors = FALSE)

# Remove final pipe delimiter
rawData$col1 <- str_replace(rawData$col1, "\\| *$", "")
head(rawData)  
```  

  
### Save out vector of column names
```{r}
colNames1 <- str_trim(unlist(str_split(line[1], "\\|")))
colNames2 <- str_trim(unlist(str_split(line[2], "\\|")))

colNames <- c(colNames1[1:2], paste(colNames1[3:10], colNames2[3:10], sep=" "))
colNames
```

  
### Split the single column into several columns based on the pipe delimiter
```{r}
splitDf <- separate(rawData, col1, into = colNames, sep = "\\|", remove = TRUE,
         convert = FALSE, extra = "warn", fill = "warn")
head(splitDf)
```

  
### Trim all columns
```{r}
splitDf <- splitDf %>% mutate_all(str_trim)
```

  
### Break this data.frame into two sets, even rows in one set, odd rows in another
```{r}
topDf <- splitDf[c(TRUE, FALSE), ] # name and rounds data
bottomDf <- splitDf[c(FALSE, TRUE), 1:2] # state and ratings data
```

  
### Parse state and ratings information
```{r}
# Add column names for reference
colnames(bottomDf) <- c("PlayerState", "temp")

# Add row id to bottomDf
bottomDf$Pair <- str_trim(topDf$Pair)

# Parse pre/post rating
bottomDf$prePostRating <- str_replace(bottomDf$temp, ".*R: ", "")
bottomDf$prePostRating <- str_replace_all(bottomDf$prePostRating, "P[:digit:]+", "")
bottomDf$preRating <- as.numeric(str_extract(bottomDf$prePostRating, "[:digit:]+"))
bottomDf$postRating <- str_extract(bottomDf$prePostRating, "[:digit:]+$")

# Select off needed columns
bottomDf <- select(bottomDf, Pair, PlayerState, preRating)

head(bottomDf)
```

  
### Parse topDf from "wide" to "long" format to deal with round information
'Points' information is parsed here for use to compare later against given "Total Pts", as a check on the work
```{r}
roundsDf <- gather(topDf, "Round", "temp", 4:10)
roundsDf$Result <- str_replace(roundsDf$temp, " .*", "")
roundsDf$Opponent <- str_trim(str_extract(roundsDf$temp, "[:digit:]+"))
roundsDf$Points <- ifelse(roundsDf$Result %in% c("W", "B", "X"), 1,
                             if_else(roundsDf$Result %in% c("D", "H"), .5,
                                     0))
head(roundsDf)
```

  
### Merge in pre-Ratings for opponents
```{r}
# merge in preRates (note... cases where there was no opponent are dropped here)
roundsDf <- left_join(roundsDf, select(bottomDf, Pair, preRating), by=c("Opponent"="Pair"))
```

  
### Calculate average opponent pre-rating, as well as a calculated total points 
```{r}
finalDf <- roundsDf %>% 
  group_by(Pair, `Player Name`, `Total Pts`) %>%
  summarize(totalPointsCalc = sum(Points, na.rm = TRUE),
            avgOppPreRating = round(mean(preRating, na.rm = TRUE), digits=0))

stopifnot(all(as.numeric(finalDf$`Total Pts`) == finalDf$totalPointsCalc))
```

  
### Merge in State and pre-rating values
```{r}
finalDf <- inner_join(bottomDf, 
                      select(finalDf, c(Pair, `Player Name`, `Total Pts`, avgOppPreRating)),
                      by="Pair")
```

  
### Clean up column names and save out .csv file
```{r}
finalDf <- finalDf %>%
  rename(PlayerName = `Player Name`, TotalPts = `Total Pts`, PlayerPreRating = `preRating`,
         AvgOppPreRating = avgOppPreRating) %>%
  select(PlayerName, PlayerState, TotalPts, PlayerPreRating, AvgOppPreRating)

head(finalDf)

write.csv(finalDf, file = "/Users/ryanweber/Desktop/CUNY/Data 607 Db/Projects/Project 1/chess.csv",
          row.names = FALSE)
```

